% Targeted for International Journal of Medical Informatics
% Word limit: 3000
%\documentclass[review]{elsarticle}
\documentclass{elsarticle}


\usepackage{lineno,hyperref,graphicx}
\usepackage{amsmath}
\usepackage[table]{xcolor}
\usepackage[section]{placeins}
\modulolinenumbers[5]
\graphicspath{ {./images/} }

% Adapted from https://www.overleaf.com/learn/how-to/Is_there_a_way_to_run_a_word_count_that_doesn%27t_include_LaTeX_commands%3F
\newcommand{\wordcount}[2]{%
  (\immediate\write18{texcount -1 -sum -merge #1.tex > #1-words.sum }%
  \input{#1-words.sum}/#2 words)%
}

\journal{Elsevier}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\newcommand{\plugin}[1]{\texttt{#1}}
\begin{frontmatter}

\title{Healthcare Pathway Discovery, Conformance, and Enrichment \wordcount{main-body}{3000}}

%% Group authors per affiliation:
%\author{Michael O’Sullivan,  Andreas W. Kempa-Liehr, Christina Lin\fnref{myfootnote}, Randall Britten, Delwyn Armstrong}
%\address{Radarweg 29, Amsterdam}
%\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
\author[mymainaddress]{Christina Lin\corref{mycorrespondingauthor}}
\author[mymainaddress]{Andreas W. Kempa-Liehr}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{a.kempa-liehr@auckland.ac.nz}

\author[Orion]{Randall Britten}
\author[Waitemata]{Delwyn Armstrong}
\author[Waitemata]{Jonathan Wallace}
\author[Waitemata]{Patrick Gladding}
\author[mymainaddress]{ Michael O'Sullivan}

\address[mymainaddress]{Department of Engineering Science, The University of Auckland, 70 Symonds St, Grafton, Auckland, New Zealand}
\address[Orion]{Orion Health, 181 Grafton Rd, Grafton, Auckland, New Zealand}
\address[Waitemata]{Waitemata District Health Board, 124 Shakespeare Rd, Takapuna, Auckland, New Zealand
}

\begin{abstract}
\subsection*{Background and purpose}
Healthcare pathways define the execution sequence of clinical activities as patients move through a treatment process, and they are critical for maintaining quality of care. The aim of this study is to investigate the utilization of business process modelling (BPM) to design an adaptive healthcare pathway mining methodology, with particular emphasis on producing pathway models that are easy to interpret for clinicians without a sufficient background in process mining.

\subsection*{Method}
This study utilizes the business process-mining software ProM to design a process mining pipeline for healthcare pathway discovery, conformance analysis, and enrichment using hospital records. The efficacy of the BPM approach is demonstrated via two case studies that apply the proposed process mining pipeline to discover appendicitis and cholecystitis 
pathways from hospital records. Machine learning methodologies based on probabilistic programming are utilised to explore pathway features that influence patient recovery time.

\subsection*{Results}
The produced appendicitis and cholecystitis 
pathway models are
easy for clinical interpretation and provides an unbiased overview of patient movements through the treatment process. Analysis of the discovered pathway model enables reasons for longer than usual treatment times to be explored and deviations from standard treatment pathways to be identified. A probabilistic regression model that estimates patient recovery time based on the information extracted by the process mining pipeline is developed and has the potential to be very useful for hospital scheduling purposes.

\subsection*{Conclusion}
This study establishes the application of the business process modelling tool ProM for the improvement of healthcare pathway mining methods. 
%There is also value in investigating the capabilities of other business process modelling tools for healthcare pathway mining purposes. 
The proposed pipeline for healthcare pathway discovery  has the potential to support the development of machine learning models to further relate healthcare pathways to performance indicators such as length of post operation stay. 

\end{abstract}

\begin{keyword}
Healthcare pathway; Process mining; Electronic health record; Propabilistic programming
%\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\input{main-body}

\section*{Author Contributions}
All authors have made substantial contribution to this study and approved the final manuscript.

\section*{Acknowledgements}
This research has been funded by the Precision Driven Health research partnership (project number 1209).

\section*{Declaration of Competing Interests}
The authors declare that they have no competing interests for this study.

\bibliography{mybibfile}

\begin{appendix}
\include{cholecystitis}
\end{appendix}
\end{document}